---
title: "Point to polygon spatial join"
output: html_notebook
---

Examples here partially adapted from one of the vignettes in the `sf` package: [3. Manipulating Simple Feature Geometries](https://r-spatial.github.io/sf/articles/sf3.html)


```{r setup, results='hide'}
# Load two necessary packages
library(tidyverse)
library(sf)
```

### First create three 2x2 polygons:
```{r make_polygons}
b0 <- st_polygon(list(rbind(c(-1,-1), c(1,-1), c(1,1), c(-1,1), c(-1,-1))))
b1 <- b0 + 2
b2 <- b0 + c(0, 2)
```

### Combine them into an sf data.frame/tibble, give them an identifier column and plot the result:
```{r polygons_sf}
polygons_sf <- st_sf(polygon_id  = LETTERS[1:3],
                     geometry = st_sfc(b0, b1, b2))

plot(polygons_sf['polygon_id'])
```

### Drop a few points in area covered by these polygons:
```{r make_points}
set.seed(100)
points_col <- st_sample(st_as_sfc(st_bbox(polygons_sf)), 15)

points_sf <- st_sf(point_id = letters[1:15],
                geometry = points_col)

ggplot() +
  geom_sf(aes(fill = polygon_id), data = polygons_sf) +
  geom_sf_label(aes(label = point_id), data = points_sf) +
  labs(title = 'Example polygons and sample points as labels')
```

### The easiest way to link polygon attributes to the point locations is using `st_join`, then output in a pretty funky plot:
```{r run_join}
points_with_polygonID <- st_join(points_sf,
                                 polygons_sf) %>% 
  mutate(point_poly_id = paste(point_id, polygon_id, sep = ' in '))

ggplot() +
  geom_sf(aes(fill = polygon_id), data = polygons_sf) +
  geom_sf_label(aes(label = point_poly_id), data = points_with_polygonID) +
  labs(title = 'Example polygons and sample points as labels')
```

